\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

I should like to express profound gratitude and appreciation to all the people who supported me during the time in which this creative work was undertaken. I especially would like to thank my parents, William and Victoria for all of their love, support and guidance. They have always encouraged me to pursue my fascination and love for music and to follow my aspirations as a musician. The earliest memories I have about music are indebted to them: since childhood they encouraged me to go to concerts and supported my piano lessons. Their intelligence, courage, kindness and commitment to social causes, continue to be a source of admiration and inspiration for me. I wish to give thanks to the rest of my family for their unconditional love and support. My brother has always supported me and has given me an example of what can be achieved through high ideals, dedication and intellectual rigor. I am grateful to my grandparents, cousins, uncles and aunts, for their emotional and intellectual encouragement, even if it is from a long distance.

This work would have not been possible without the encouragement and guidance of my supervisors Richard Barrett and Christopher Fox. They have taught me the importance of  critical reflection in music and the virtue and integrity in attempting to create original and challenging music with diligent care and attention. They have both given me thoughtful comments and suggestions and their interest and enthusiasm for my work has been a constant source of encouragement and motivation. I would particularly like to thank Chris, for giving me advice and support while writing the commentary and for having been an invaluable mentor for these four years.

I am grateful for the financial support of the Overseas Research Scheme (ORS) and Brunel University School of Arts, which made my study in the United Kingdom possible. The PRS for Music Foundation gave financial assistance for the performance of my music.

During this time, I have been extremely fortunate to collaborate with musicians of the highest level. I would especially like to thank Javier Carmona, Adam de la Cour, Aleksander Kolkowski, Dominic Lash, Alexander Hawkins, Neil Luck and Nicholas Peters for their musicianship and friendship. I have learned a lot by collaborating with them and I have enjoyed their company as much as making music with them. I would also like to thank \textbf{piano}\emph{circus} and Sarah Nicolls for the time and effort they devoted to playing my compositions and for dealing with the unusual tasks involved in performing them.
 
Steve Potter has been a friend, colleague and neighbor who has given me crucial support, inspiration, advice and criticism and as a performer has championed my work. I have also benefited from conversations and insights from Ed McKeon, who has also helped promote and produce performances of my music. Steve, Ed and Philip Somervell have shared my interest in critical theory, aesthetics, and philosophy and through various discussions they provided useful comments that influenced this written commentary. I would also like to thank my fellow composers with whom I have shared conversations and musical experiences that have informed and inspired my work in significant ways. In particular, I thank Jasna Veli\v{c}kovi\'{c}, Ronald Boersen, Dganit Elyakim, Assaf Gidron, Ophir Ilzetzki, Justin Christensen and Mauricio Pauly. I extend my gratitude to my previous composition teachers: Louis Andriessen, Richard Ayres, Gilius van Bergeijk, Martijn Padding and Lawrence Moss. They have been responsible for much of the musical knowledge I have gained and the aesthetic direction I have chosen and I am indebted to them.

I should acknowledge the SuperCollider community, not only for pointing me in the right direction as I was learning how to use this great tool, but also for continuously developing it as an open source project. David Plans Casal and Gabriel Montagn\'e gave me great computer programming tips, explained useful programming concepts and encouraged me to further my interest in computer music.

Finally, I would like to thank Eszter, who came together with me to London and has been at my side during this entire period. She has given me more love and support than I could have ever hoped for. Together we have gone through the ups and downs of living in London and she has supported me in the most trying moments of this research period. With her I have shared many moments and laughs and she has also inspired me through her passion for life and her multiple interests and talents in art, cooking, photography, design, decoration, crafts and fashion, to mention just a few. She has brought me happiness and strength through her sweetness, comfort and loving company during the past four years.  
