%\begin{thebibliography}{99} 
%\chapter{Bibliography}
\paragraph{}
\paragraph{}
\begin{flushleft}
\huge 
\marginpar{}
\title{\textbf{Bibliography}} \\
\end{flushleft}
\vspace{20pt} 
\normalsize
Aristotle (1995), \emph{Politics}, Trans. Ernest Barker, Oxford: Oxford University Press.\\
Babbitt, Milton (2003), `The Composer as Specialist,' in \emph{The Collected Essays of Milton Babbitt}.  

ed. Stephen Peles,  Princeton: Princeton University Press.\\
Badiou, Alain (2001), `The Ethics of Truths,' in \emph{Ethics: An Essay on the Understanding of Evil}, 

Trans. Peter Hallward, London: Verso.\\
------ (2006), \emph{Infinite Thought}, Trans. Oliver Feltham and Justin Clemens, London: Continuum.\\
------ (2009), `Scholium: A Musical Variant of the Metaphysics of the Subject,' in \emph{Logics of Worlds:}

\emph{Being and Event II}, Trans. Alberto Toscano, London: Continuum.\\
Barlow, Clarence (2000), \emph{Musica Derivata} [CD], hat[now]ART 126, Hat Hut Records Basle.\\
Berio, Luciano (1985), \emph{Two Interviews with Rossana Dalmonte and B\'{a}lint Andr\'{a}s Varga},  Ed. and 

Trans. David Osmond-Smith, London: Marion Bowars.\\
Born, Georgina (1995), \emph{Rationalizing Culture : IRCAM, Boulez, and the Institutionalization of the}

\emph{Musical Avant-Garde}, Berkeley: University of California Press.\\
Boulez, Pierre (1991), `Sch\"{o}nberg is dead,' in \emph{Stocktakings from an Apprenticeship}, Oxford: Oxford 

University Press.\\
Bourriaud, Nicolas (2005), \emph{Postproduction. Culture as Screenplay: How Art Reprograms the World}, 

New York: Lukas and Sternberg.\\
Emmerson, Simon (2000),`Crossing cultural boundaries through technology?,'  in \emph{Music, Electronic}

\emph{Media and Culture}, Aldershot: Ashgate.\\
Ferneyhough, Brian (1995), `Aspects of Notational and Compositional Practice,' in \emph{Collected}

\emph{Writings}, ed. James Boros and Richard Toop, Amsterdam: Harwood Academic Publishers.\\
Goehr, Lydia (2007), \emph{The Imaginary Museum of Musical Works: An Essay in the Philosophy of}

\emph{Music}, Oxford: Oxford University Press.\\
Levitin, Daniel (2006), \emph{This is Your Brain on Music: Understanding a Human Obsession}, London: 

Atlantic Books.\\
Lyotard, Jean-Fran\c{c}ois (1993), `To Mathias Khan,' in \emph{The Postmodern Explained to Children:} 

\emph{Correspondance 1982-1985}, Trans. Julian Pefanis and Morgan Thomas, London: Turnaround.\\
Metzer, David (2003), \emph{Quotation and Cultural Meaning in Twentieth-Centure Music}, Cambridge: 

Cambridge University Press.\\
Mithen, Steven (2006), \emph{The Singing Neanderthals: The Origins of Music, Language, Mind and Body}, 

London: Phoenix.\\
Nono, Luigi (1975), \emph{Texte. Studien zu seiner Musik}, ed. J\"{u}rg Stenzl, Z\"{u}rich: Atlantis.\\
Osmond-Smith, David (1985), \emph{Playing with Words: a Guide to Luciano Berio's Sinfonia},  London: 

The Royal Musical Association.\\
Oswald, John (1985),`Plunderphonics, or Audio Piracy as a Compositional Prerogative,' in \emph{Wired}

\emph{Society Electro-Acoustic Conference}, Toronto, 1985. URL: \texttt{http://www.plunderphonics.com/xht}

\texttt{ml/xplunder.html.}\\
Plato (2006), \emph{The Republic}, Trans. R.E. Allen, New Haven: Yale University Press.\\
Pinker, Steven (1998), \emph{How the Mind Works}, London: Allen Lane The Penguin Press.\\
Ranci\`{e}re, Jaques (2004), \emph{The Politics of Aesthetics}, Trans. Gabriel Rockhill, London: Continuum. \\
------ (2007), \emph{The Future of the Image}, Trans. Gregory Elliot, London: Verso.\\
------ (2009), \emph{Aesthetics and Its Discontents}, Trans. Steven Corcoran, Cambridge: Polity.\\
Said, Edward W. (1992), \emph{Musical Elaborations}, London: Vintage.\\
Sch\"{o}nberg, Arnold (1984),  \emph{Style and Idea}, Trans. Leo Black, Los Angeles: University of California

Press.\\
Small, Christopher (1998), \emph{Musicking: The Meanings of Performing and Listening}, Middletown, 

Connecticut: Wesleyan University Press.\\
Taruskin, Richard (2009), \emph{The Danger of Music}, Berkley: University of California Press.\\
\v{Z}i\v{z}ek, Slavoj (2006), \emph{The Universal Exception}, London: Continuum.\\\
------ (2009), \emph{First As Tragedy, Then as Farce}, London: Verso.\\
%\end{thebibliography} 
