%\begin{thebibliography}{99} 
%\chapter{Bibliography}
\paragraph{}
\paragraph{}
\begin{flushleft}
\huge 
\marginpar{}
\textbf{Bibliography}
\end{flushleft}
\vspace{20pt} 
\normalsize
\hypertarget{adornointro}{}
Adorno, Theodor W. (1976), \emph{Introduction to the Sociology of Music}, trans. E. B. Ashton, New York: 

The Seabury Press. 
\hypertarget{adornoaesth}{}\\
------ (1997), \emph{Aesthetic Theory}, trans. Robert Hullot-Kentor, Minneapolis: University of Minnesota 

Press. 
\hypertarget{aristotle}{}\\
Aristotle (1995), \emph{Politics}, trans. Ernest Barker, Oxford: Oxford University Press. 
\hypertarget{babbitt}{}\\
Babbitt, Milton (2003), `The Composer as Specialist', in \emph{The Collected Essays of Milton Babbitt}.  

ed. Stephen Peles,  Princeton: Princeton University Press, pp. 48-54.
\hypertarget{badiouethics}{}\\
Badiou, Alain (2001), `The Ethics of Truths', in \emph{Ethics: An Essay on the Understanding of Evil}, 

trans. Peter Hallward, London: Verso. pp. 40-57.
\hypertarget{infthought}{}\\
------ (2006), \emph{Infinite Thought}, trans. Oliver Feltham and Justin Clemens, London: Continuum. 
\hypertarget{badioumus}{}\\
------ (2009), `Scholium: A Musical Variant of the Metaphysics of the Subject', in \emph{Logics of Worlds:}

\emph{Being and Event II}, trans. Alberto Toscano, London: Continuum, pp. 79-89. 
\hypertarget{barlowspec}{}\\
Barlow, Clarence (1998), `On the spectral analysis of speech for subsequent resynthesis by acoustic 

instruments', in \emph{Forum phoneticum}, vol. 66, pp. 183-190.
\hypertarget{barlowcd}{}\\
------ (2000), \emph{Musica Derivata} [CD], hat[now]ART 126, Basel: Hat Hut Records. 
\hypertarget{barlowbob}{}\\
------ (2007), `Clarence Barlow: Interview by Bob Gilmore', in \emph{PARIS Transatlantic}, Paris, 2007, 

URL: \href{http://www.paristransatlantic.com/magazine/interviews/barlow.html}{\texttt {http://www.paristransatlantic.com/magazine/interviews/barlow.html}}.
\hypertarget{barlowpiano}{}\\
------ (2008), \emph{Klavierwerke von Klarenz Barlow} [CD], Feedback CD8, D\"{u}sseldorf: Cybele Records. 
\hypertarget{barthes}{}\\
Barthes, Roland (1977), `Music Practica,' in \emph{Image Music Text}, trans. Stephen Heath, London: 

Fontana Press, pp. 149-154. 
\hypertarget{berio}{}\\
Berio, Luciano (1985), \emph{Two Interviews with Rossana Dalmonte and B\'{a}lint Andr\'{a}s Varga},  ed. and 

trans. David Osmond-Smith, London: Marion Bowars. 
\hypertarget{born}{}\\
Born, Georgina (1995), \emph{Rationalizing Culture : IRCAM, Boulez, and the Institutionalization of the}

\emph{Musical Avant-Garde}, Berkeley: University of California Press. 
\hypertarget{boulez}{}\\
Boulez, Pierre (1991), `Sch\"{o}nberg is dead', in \emph{Stocktakings from an Apprenticeship}, Oxford: Oxford 

University Press, pp. 212-213??? 
\hypertarget{relational}{}\\
Bourriaud, Nicolas (2002), \emph{Relational Aesthetics}, trans. Pleasance Simon and Fronza Woods, Dijon: 

Les presses du r\'{e}el.
\hypertarget{postproduction}{}\\
------ (2005),  \emph{Postproduction. Culture as Screenplay: How Art Reprograms the World}, 

New York: Lukas and Sternberg.
\hypertarget{cage}{}\\
Cage, John (1961), \emph{Imaginary Landscape No. 5} [Music Score], New York: Henmar Press. 
\hypertarget{cutler}{}\\
Cutler, Chris (2006),  `Plunderphonia', in \emph{Audio Culture}, eds. Christoph Cox and Daniel Warner, 

London: Continuum, pp. 138-156.
\hypertarget{debord}{}\\
Debord, Guy (1994), \emph{The Society of the Spectacle}, trans. Donald Nicholson-Smith, New York: Zone 

Books.
\hypertarget{debord2}{}\\
------ (2009), `The Use of Stolen Films', in \emph{Appropriation}, ed. David Evans, trans. Ken Knabb. 

London: Whitechapel, p. 66.
\hypertarget{debord3}{}\\
Debord, Guy and Gil J. Wolman (2009), `Directions for the Use of D\'{e}tournement', in \emph{Appropriation}, 

ed. David Evans, trans. Ken Knabb. London: Whitechapel, pp. 35-39.
\hypertarget{duchamp}{}\\
Duchamp, Marcel (2009), `Apropos of `Readymades'', in \emph{Appropriation}, ed. David Evans, London: 

Whitechapel, p. 40.
\hypertarget{emmersoncross}{}\\
Emmerson, Simon (2000),`Crossing cultural boundaries through technology?',  in \emph{Music, Electronic}

\emph{Media and Culture}, Aldershot: Ashgate, pp. 115-137. 
\hypertarget{emmersonliving}{}\\
------ (2007), \emph{Living Electronic Music}, Surrey: Ashgate. 
\hypertarget{eno}{}\\
Eno, Brian (1996), `Generative Music: Evolving metaphors, in my opinion, is what artists do', in \emph{In}

\emph{Motion Magazine}, San Francisco, 1996, URL: \href{http://www.inmotionmagazine.com/eno1.html}{\texttt {http://www.inmotionmagazine.com/eno1.html}}.
\hypertarget{evans}{}\\
Evans, David (2009), `Seven Types of Appropriation', in \emph{Appropriation}, ed. David Evans, London: 

Whitechapel, pp. 12-23.
\hypertarget{ferneyhough}{}\\
Ferneyhough, Brian (1995), `Aspects of Notational and Compositional Practice', in \emph{Collected}

\emph{Writings}, ed. James Boros and Richard Toop, Amsterdam: Harwood, pp. 2-13.
\hypertarget{ford}{}\\
Ford, Simon (2005), \emph{The Situationist International: A User's Guide}, London: Black Dog Publishing.
\hypertarget{fox}{}\\
Fox, Christopher (2007), `Where the river bends: the Cologne School in retrospect', in  \emph{Musical Times}, 

Winter 2007, pp. 27-42.
\hypertarget{goer}{}\\
Goehr, Lydia (2007), \emph{The Imaginary Museum of Musical Works: An Essay in the Philosophy of}

\emph{Music}, Oxford: Oxford University Press.
\hypertarget{hamilton}{}\\
Hamilton, Andy (2008), \emph{Aesthetics and Music}, London: Continuum. 
 \hypertarget{hallward}{}\\
Hallward, Peter (2008), `Order and event: on Badiou's Logics of worlds'. New Left Review, vol. 53, 

pp. 97-122, URL: \href{http://nsrnicek.googlepages.com/hallward_order_and_event.pdf}{\texttt{http://nsrnicek.googlepages.com/hallward\_order\_and\_event.pdf}}.
\hypertarget{judovitz}{}\\
Judovitz, Dalia (1995), \emph{Unpacking Duchamp: art in transit}, Berkley: University of California Press. 
\hypertarget{klingbeil}{}\\
Klingbeil, Michael (2005), `Software for spectral analysis, editing, and synthesis', in \emph{Proceedings of}

\emph{ICMC}, vol. 2005, URL: \href{http://www.klingbeil.com/papers/spearfinal05.pdf}{\texttt{http://www.klingbeil.com/papers/spearfinal05.pdf}}.
\hypertarget{musmind}{} \\ 
Levitin, Daniel (2006), \emph{This is Your Brain on Music: Understanding a Human Obsession}, London: 

Atlantic Books. 
\hypertarget{lyotard}{}\\
Lyotard, Jean-Fran\c{c}ois (1992), `Missive on Universal History', in \emph{The Postmodern Explained to}

\emph{Children:} \emph{Correspondance 1982-1985}, trans. Julian Pefanis and Morgan Thomas, London: 

Turnaround, pp. 35-47. 
\hypertarget{mccartney}{}\\
McCartney, James (2002), `Rethinking the Computer Language: SuperCollider', in \emph{Computer Music} 

\emph{Journal}, volume 25, number 4.
\hypertarget{metzer}{}\\
Metzer, David (2003), \emph{Quotation and Cultural Meaning in Twentieth-Centure Music}, Cambridge: 
\hypertarget{mcdonough}{}\\
McDonough, Tom, ed. (2002), \emph{Guy Debord and The Situationist International: Texts and Documents}, 

London: MIT Press. 
\hypertarget{mithen}{}\\
Mithen, Steven (2006), \emph{The Singing Neanderthals: The Origins of Music, Language, Mind and Body}, 

London: Phoenix. 
\hypertarget{moholy}{}\\
Moholy-Nagy, L\'{a}szl\'{o} (2006),  `Production-Reproduction: Potentialities of the Phonograph', in \emph{Audio} 

\emph{Culture}, eds. Christoph Cox and Daniel Warner, London: Continuum, pp. 331-333.
\hypertarget{nono}{}\\
Nono, Luigi (1975), \emph{Texte. Studien zu seiner Musik}, ed. J\"{u}rg Stenzl, Z\"{u}rich: Atlantis. 
\hypertarget{ossmith}{}\\
Osmond-Smith, David (1985), \emph{Playing with Words: a Guide to Luciano Berio's Sinfonia},  London: 

The Royal Musical Association. 
\hypertarget{oswald}{}\\
Oswald, John (1985),`Plunderphonics, or Audio Piracy as a Compositional Prerogative', in \emph{Wired}

\emph{Society Electro-Acoustic Conference}, Toronto, 1985. URL: \href{http://www.plunderphonics.com/xhtml/xplunder.html}{\texttt{http://www.plunderphonics.com/xht}}

\href{http://www.plunderphonics.com/xhtml/xplunder.html}{\texttt{ml/xplunder.html}}.
\hypertarget{pfaller}{}\\
Pfaller, Robert (2003), `Backup of Little Gestures of Disappearance: Interpassivity and the Theory 

of Ritual' in \emph{Journal of European Psychoanalysis}, JEP Online, number 16, Winter-Spring 2003, 

URL: \href{http://www.psychomedia.it/jep/number16/pfaller.htm}{\texttt {http://www.psychomedia.it/jep/number16/pfaller.htm}}. 
\hypertarget{plato}{}\\
Plato (2006), \emph{The Republic}, trans. R.E. Allen, New Haven: Yale University Press. 
\hypertarget{pinker}{}\\
Pinker, Steven (1998), \emph{How the Mind Works}, London: Allen Lane The Penguin Press. 
\hypertarget{ranpoli}{}\\
Ranci\`{e}re, Jaques (2004), \emph{The Politics of Aesthetics}, trans. Gabriel Rockhill, London: Continuum. 
\hypertarget{ranimg}{}\\
------ (2007), \emph{The Future of the Image}, trans. Gregory Elliot, London: Verso. 
\hypertarget{ranaesth}{}\\
------ (2009), \emph{Aesthetics and Its Discontents}, trans. Steven Corcoran, Cambridge: Polity. 
\hypertarget{ranspec}{}\\
------ (2009), \emph{The Emancipated Spectator}, trans. Gregory Elliot, London: Verso.
\hypertarget{reuben}{}\\
Reuben, Federico and Aleksander Kolkowski (2008), `Horatio Oratorio: composing using historic 

sound recordings', in \emph{Proceedings of CHARM}, Symposium 6, 2008, URL: \href{http://images.cch.kcl.ac.uk/charm/liv/redist/pdf/s6kolkowski_reuben.pdf}{\texttt{http://images.cch.kcl.}}

\href{http://images.cch.kcl.ac.uk/charm/liv/redist/pdf/s6kolkowski_reuben.pdf}{\texttt{ac.uk/charm/liv/redist/pdf/s6kolkowski\_reuben.pdf}}.
\hypertarget{rowe}{}\\
Rowe, Robert (2001), \emph{Machine Musicianship}, Cambridge: MIT Press. 
\hypertarget{said}{}\\
Said, Edward W. (1992), \emph{Musical Elaborations}, London: Vintage. 
\hypertarget{schoen}{}\\
Sch\"{o}nberg, Arnold (1984),  \emph{Style and Idea}, trans. Leo Black, Los Angeles: University of California

Press. 
\hypertarget{small}{}\\
Small, Christopher (1998), \emph{Musicking: The Meanings of Performing and Listening}, Middletown, 

Connecticut: Wesleyan University Press. 
\hypertarget{stelarc}{}\\
Stelarc (2004), `Zombies and Cyborgs: The Cadaver, the Comatose and the Chimera', London, 2004.

URL: \href{http://www.stelarc.va.com.au/documents/zombiesandcyborgs.pdf}{\texttt{http://www.stelarc.va.com.au/documents/zombiesandcyborgs.pdf}}.
\hypertarget{taruskin}{}\\
Taruskin, Richard (2009), \emph{The Danger of Music}, Berkley: University of California Press. 
\hypertarget{wishart}{}\\
Wishart, Trevor (1996), \emph{On Sonic Art}, London: Routledge. 
\hypertarget{xenakis}{}\\
Xenakis, Iannis (1992), \emph{Formalized Music: Thought and Mathematics in Music}, Hillsdale, NY: 

Pendragon Press. 
\hypertarget{zizekuniv}{}\\
\v{Z}i\v{z}ek, Slavoj (2006), \emph{The Universal Exception}, London: Continuum. 
\hypertarget{zizekreader}{}\\\
------ (2006), \emph{The \v{Z}i\v{z}ek Reader}, Oxford: Blackwell.
\hypertarget{zizekviolence}{}\\
------ (2008), \emph{Violence}, London: Profile Books. 
\hypertarget{zizektragedy}{}\\
------ (2009), \emph{First As Tragedy, Then as Farce}, London: Verso.\\
%\end{thebibliography} 
